---
title: "2a. CM_wmhv_02 24 25"
output: html_document
date: "2025-02-24"
---
# Notes 
This example code below generates exposure and mediator (binary: hypertension) IPWs for the outcome of WMHV. Additionally, we estimate total and controlled directed effects using the survey package. Similar code was used to model AD meta-ROI cortical thickness, and amyloid and tau deposition. We also summarize the weights here. 

Some minor changes: 
1) For amyloid and tau outcomes, APOE4 status was added to the exposure models.
2) PET scanner was added to outcome models for amyloid - note that with our data, there was no variability in the PET scanner for tau, so that variable was excluded from the model. 

# Getting packages into session
```{r packages, include = FALSE}
knitr::opts_chunk$set(echo = TRUE, results = "hide")

library(haven) # inputting data 
library(tidyverse) # data mgmt
library(psych) # easy summary statistics
library(DT) # data tables
library(tableone) # easy table 1 
library(kableExtra) # format kable objects 
library(naniar) # for missingness 
library(survey) # for MSMs 
library(nnet) # multinormal regression
library(cobalt) # examining covariate balance
library(gt) # nice tables
library(gtExtras) # nice tables
library(gtsummary) # nice summary tables
library(broom) # making output nice
library(mice) # working with imputed data
options(max.print = 100000)
options(scipen = 999) # turn off scientific notation 
set.cobalt.options(binary='raw', continuous='std') # settings for checking covariate balance 
```
```{css scroll box for code, include = FALSE}
pre {
  max-height: 300px;
  overflow-y: auto;
}

pre[class] {
  max-height: 100px;
}
```
# Uploading data
```{r upload data, include = FALSE}
load(file = "./analysis data/causmed_imp.Rdata") 
view(names(causmed_imp[[1]])) # getting names for reference

# creating dataset with imputed datasets stacked together
imp_stacked <- complete(causmed_imp, "long") # n=71840, v=68, 2 additional variables to label imputations + id 
str(imp_stacked)
summary(imp_stacked$".imp")
table(imp_stacked$".imp")
```
# Exposure weights
This accounts for exposure-outcome confounding (race/eth -> neuroimg). We estimate the probability of being Black or Hispanic (numerator), then the probability of being Black or Hispanic conditional on confounders of interest. We will use age and gender as confounders here since there are very few true confounders of the E-O relationship (many are mediators). 

References:
- https://pmc.ncbi.nlm.nih.gov/articles/PMC3553264/
- https://pubmed.ncbi.nlm.nih.gov/36938776/
- https://pmc.ncbi.nlm.nih.gov/articles/PMC3710547

## Numerator model for exposure
```{r num exp, echo = TRUE}
# Prepping some lists/variables for for-loop 
m <- max(imp_stacked$".imp") # number of imputed datasets
num_race_wmhv <- list() # for fitted values
num_race <- list() # numerator for each race/ethnicity 

# For-loop to obtain numerator model for exposure 
for (i in seq_along(1:m)){
  
  # extract one imputed dataset at a time
  data <- imp_stacked %>% filter(`.imp` == i)
  
  # multinomial regression to obtain unconditional prob 
  # of being in a race/ethnic group
  num_race_wmhv[[i]] <- multinom(as.factor(ethnicity) ~ 1,
                                 data = data)
  
  # obtaining fitted values 
  num_race[[i]] <- num_race_wmhv[[i]]$fitted.values %>%
    as.data.frame() %>%
    rename(num_white = White,
           num_black = Black,
           num_hispanic = Hispanic)
  
}

# checking probabilities
str(num_race)
summary(num_race[[20]]) # switch index to check all datasets

# cleaning up workspace for next section
rm(i, m, data, num_race_wmhv)

```
## Denominator model for exposure
```{r den exp, echo = TRUE}
# Prepping some lists/variables for for-loop 
m <- max(imp_stacked$".imp") # number of imputed datasets
den_race_wmhv <- list() # for fitted values
den_race <- list() # numerator for each race/ethnicity 

# For-loop to obtain numerator model for exposure 
for (i in seq_along(1:m)){
  
  # extract one imputed dataset at a time
  data <- imp_stacked %>% filter(`.imp` == i)
  
  # multinomial regression to obtain conditional prob 
  # of being in a race/ethnic group
  den_race_wmhv[[i]] <- multinom(as.factor(ethnicity) ~ 
                            age_1 + # using visit 1 age here 
                            as.factor(gender),
                          data = data)
  
  # obtaining fitted values 
  den_race[[i]] <- den_race_wmhv[[i]]$fitted.values %>%
    as.data.frame() %>%
    rename(den_white = White,
           den_black = Black,
           den_hispanic = Hispanic)
  
}

# checking probabilities
str(den_race)
summary(den_race[[20]]) # switch index to check all datasets

# cleaning up workspace for next section
rm(i, m, data, den_race_wmhv)
```
## Calculating exposure weights
```{r exp IPW, echo = TRUE}
# Prepping some lists/variables for for-loop 
m <- max(imp_stacked$".imp") # number of imputed datasets
imp_stacked_exp <- list() # list for exposure weights 

# For-loop to obtain numerator model for exposure 
for (i in seq_along(1:m)){
  
  # extract one imputed dataset at a time
  data <- imp_stacked %>% filter(`.imp` == i)
  
  # Creating stabilized IPW for exposure
  imp_stacked_exp[[i]] <- data %>%
  
  # getting probabilities from the models above
  bind_cols(num_race[[i]],
            den_race[[i]]) %>% 

  # creating denominators based on exposure status
  mutate(num_exp = case_when(
    ethnicity == "Black" ~ num_black,
    ethnicity == "Hispanic" ~ num_hispanic,
    ethnicity == "White" ~ num_white,
    TRUE ~ NA)) %>%
  
  # creating denominators based on exposure status
  mutate(den_exp = case_when(
    ethnicity == "Black" ~ den_black,
    ethnicity == "Hispanic" ~ den_hispanic,
    ethnicity == "White" ~ den_white,
    TRUE ~ NA)) %>% 
  
  # creating exposure weights
  mutate(sw_exp = (num_exp/den_exp)) 
  
}
```
## Examining exposure IPWs
Weights have a mean ~1 and none >10. 
```{r check IPW exp, echo = TRUE}
# switch index to check each dataset separately 
summary(imp_stacked_exp[[5]]$num_exp) 
describeBy(imp_stacked_exp[[5]]$num_exp, 
           imp_stacked_exp[[5]]$ethnicity, 
           mat = TRUE) 

summary(imp_stacked_exp[[5]]$den_exp)
describeBy(imp_stacked_exp[[5]]$den_exp, 
           imp_stacked_exp[[5]]$ethnicity, 
           mat = TRUE) 

summary(imp_stacked_exp[[20]]$sw_exp) 
describeBy(imp_stacked_exp[[2]]$sw_exp, 
           imp_stacked_exp[[2]]$ethnicity, 
           mat = TRUE) 

bal.tab(ethnicity ~ age_1 + gender,
        data = imp_stacked_exp[[20]],
        weights = "sw_exp")

# cleaning up work space
rm(i, m, num_race, den_race)
```
# Mediator weights
This accounts for mediator-outcome confounding (HTN -> neuroimg). We estimate the probability of having hypertension, then the probability of having hypertension conditional on M-O confounders (age, gender, SES markers + physical activity + smoking + BMI). 

## Numerator model for mediator
```{r num med, echo = TRUE}
# Prepping some lists/variables for for-loop 
m <- max(imp_stacked$".imp") # number of imputed datasets
num_sbp_wmhv <- list() # for fitted values
num_med <- list() # numerator for sbp 

# For-loop to obtain numerator model for exposure 
for (i in seq_along(1:m)){
  
  # extract one imputed dataset at a time
  data <- imp_stacked %>% filter(`.imp` == i)
  
  # logistic regression to obtain unconditional prob 
  # of being in a race/ethnic group
  num_sbp_wmhv[[i]] <- lm(sbpavg ~ 1, data = data)
  
  # obtaining probability density function
  num_med[[i]] <- dnorm(x = data$sbpavg,
                        mean = predict(num_sbp_wmhv[[i]]),
                        sd = sd(num_sbp_wmhv[[i]]$residuals)) %>% # SD of residuals 
    as.data.frame() %>%
  rename(num_med = ".")
}

# checking probabilities
str(num_med)
summary(num_med[[20]]) # switch index to check all datasets

# cleaning up workspace for next section
rm(i, m, data, num_sbp_wmhv)
```
## Denominator model for mediator
```{r den med, echo = TRUE}
# Prepping some lists/variables for for-loop 
m <- max(imp_stacked$".imp") # number of imputed datasets
den_sbp_wmhv <- list() # for fitted values
den_med <- list() # denominator for htn

# For-loop to obtain numerator model for exposure 
for (i in seq_along(1:m)){
  
  # extract one imputed dataset at a time
  data <- imp_stacked %>% filter(`.imp` == i)
  
  # linear regression to obtain conditional probability 
  den_sbp_wmhv[[i]] <- lm(sbpavg ~
                            as.factor(ethnicity) +
                            age_1 + # using baseline age 
                            as.factor(gender) +
                            # SES variables 
                            income +
                            as.factor(hasnoinsurance) + 
                            # proximal risk factors
                            bmi +
                            rapa_1_total + # physical act scores
                            rapa_2_total +
                            as.factor(smkever) +  
                            # taking BP meds to continuous model
                            # dbpavg,
                            as.factor(takes_BP_meds),
                          data = data)
  
  # obtaining probability distribution functions 
  den_med[[i]] <- dnorm(x = data$sbpavg, 
                        mean = predict(den_sbp_wmhv[[i]]),
                        sd = sd(den_sbp_wmhv[[i]]$residuals)) %>% # SD of residuals
    as.data.frame() %>%
    rename(den_med = ".")
}

# checking probabilities
str(den_med)
summary(den_med[[20]]) # switch index to check all datasets

# cleaning up workspace for next section
rm(i, m, data, den_sbp_wmhv)
```
## Calculating mediator weights
```{r med IPW, echo = TRUE}
# Prepping some lists/variables for for-loop 
m <- max(imp_stacked$".imp") # number of imputed datasets
imp_stacked_med <- list() # list for mediator weights 

# For-loop to obtain numerator model for exposure 
for (i in seq_along(1:m)){
  
  # extract one imputed dataset at a time
  data <- imp_stacked %>% filter(`.imp` == i)
  
  # Creating stabilized IPW for exposure
  imp_stacked_med[[i]] <- data %>%
  
  # getting probabilities from the models above
  bind_cols(num_med[[i]], 
            den_med[[i]]) %>% 
    
  # creating mediator weights
  mutate(sw_med = (num_med/den_med))
}

```
## Examining mediator IPWs
```{r check IPW med, echo = TRUE}
# switch index to check each dataset separately 
summary(imp_stacked_med[[5]]$num_med)
describeBy(imp_stacked_med[[5]]$num_med, 
           imp_stacked_med[[5]]$has_htn, 
           mat = TRUE) 

summary(imp_stacked_med[[1]]$den_med)
describeBy(imp_stacked_med[[1]]$den_med, 
           imp_stacked_med[[1]]$has_htn, 
           mat = TRUE) 

summary(imp_stacked_med[[20]]$sw_med) 
describeBy(imp_stacked_med[[20]]$sw_med, 
           imp_stacked_med[[20]]$has_htn, 
           mat = TRUE) 

bal.tab(has_htn ~ 
          ethnicity +
          age_1 + # using baseline age 
          gender +
          # SES variables 
          edu + 
          income +
          hasnoinsurance + 
          # proximal risk factors
          bmi +
          rapa_1_total + # physical act scores
          rapa_2_total + 
          smkever,
        data = imp_stacked_med[[5]],
        weights = "sw_med")

# cleaning up work space
rm(i, m, data, num_med, den_med)
```
# Selection weights 
This accounts for selection into the imaging cohort, in this case, who got an MRI and had WMHV measured. In theory, this would be the same as the AD-meta ROI CT measure, but we will estimate separately for each outcome because N's are still slightly different even for same imaging modality. Note that we are using 1 attrition weight, vs. using 1 for attrition and 1 for death (something to consider for the future). 

## Numerator model for selection
```{r num sel, echo = TRUE}
# Prepping some lists/variables for for-loop 
m <- max(imp_stacked$".imp") # number of imputed datasets
sel_wmhv <- list() # for fitted values
num_sel_wmhv <- list() # numerator for selection into the subsample 

# For-loop to obtain numerator model for exposure 
for (i in seq_along(1:m)){
  
  # extract one imputed dataset at a time
  data <- imp_stacked %>% filter(`.imp` == i)
  
  # logistic regression, prob of being in the WMHV sample 
  sel_wmhv[[i]] <- glm(as.factor(haswmhv2) ~ 1,
                       family = binomial(link = "logit"),
                       data = data)
  
  # obtaining fitted values 
  num_sel_wmhv[[i]] <- sel_wmhv[[i]]$fitted.values %>%
    as.data.frame() %>%
  rename(num_sel = ".")
}

# checking probabilities
str(num_sel_wmhv)
summary(num_sel_wmhv[[20]]) # switch index to check all datasets

# cleaning up workspace for next section
rm(i, m, data, sel_wmhv)
```
## Denominator model for selection 
```{r den sel, echo = TRUE}
# Prepping some lists/variables for for-loop 
m <- max(imp_stacked$".imp") # number of imputed datasets
sel_wmhv <- list() # for fitted values
den_sel_wmhv <- list() # numerator for selection into the subsample 

# For-loop to obtain numerator model for exposure 
for (i in seq_along(1:m)){
  
  # extract one imputed dataset at a time
  data <- imp_stacked %>% filter(`.imp` == i)
  
  # logistic regression, prob of being in the WMHV sample 
  sel_wmhv[[i]] <- glm(as.factor(haswmhv2) ~ 
                         as.factor(ethnicity) +
                         age_1 + # using baseline age 
                         as.factor(gender) +
                         # SES variables 
                         edu + 
                         income +
                         as.factor(hasnoinsurance) + 
                         socsupptot + 
                         # proximal risk factors
                         bmi +
                         rapa_1_total + # physical act scores
                         rapa_2_total + 
                         as.factor(smkever) + 
                         gds_total + 
                         # lab values for cardiometabolic RF
                         ldl + 
                         gluc + 
                         a1c + 
                         egfr_nonaa + 
                         egfr_aa + 
                         choltot + 
                         # clinical dx vars
                         as.factor(cdx_dep) +
                         as.factor(cdx_dyslipid) + 
                         as.factor(cdx_cvd) + 
                         as.factor(cdx_dm), 
                       family = binomial(link = "logit"),
                       data = data)
  
  # obtaining fitted values 
  den_sel_wmhv[[i]] <- sel_wmhv[[i]]$fitted.values %>%
    as.data.frame() %>%
  rename(den_sel = ".")
}

# checking probabilities
str(den_sel_wmhv)
summary(den_sel_wmhv[[20]]) # switch index to check all datasets

# cleaning up workspace for next section
rm(i, m, data, sel_wmhv)
```
## Calculating selection weights
```{r sel IPW, echo = TRUE}
# Prepping some lists/variables for for-loop 
m <- max(imp_stacked$".imp") # number of imputed datasets
imp_stacked_sel <- list() # list for selection weights 

# For-loop to obtain numerator model for exposure 
for (i in seq_along(1:m)){
  
  # extract one imputed dataset at a time
  data <- imp_stacked %>% filter(`.imp` == i)
  
  # Creating stabilized IPW for exposure
  imp_stacked_sel[[i]] <- data %>%
  
  # getting probabilities from the models above
  bind_cols(num_sel_wmhv[[i]], 
            den_sel_wmhv[[i]]) %>% 

  # creating numerators based on mediator status
  mutate(num_sel_f = case_when(
    haswmhv2 == 1 ~ num_sel,
    haswmhv2 == 0 ~ 1-num_sel,
    TRUE ~ NA)) %>%
  
  # creating denominators based on mediator status
  mutate(den_sel_f = case_when(
    haswmhv2 == 1 ~ den_sel,
    haswmhv2 == 0 ~ 1-den_sel,
    TRUE ~ NA)) %>%
  
  # creating mediator weights
  mutate(sw_sel = (num_sel_f/den_sel_f))
}
```
## Examining selection IPWs
Weights mean ~ 1, max 11.9, will consider truncating. 
```{r check sel med, echo = TRUE}
# switch index to check each dataset separately 
summary(imp_stacked_sel[[5]]$num_sel_f)
describeBy(imp_stacked_sel[[5]]$num_sel_f, 
           imp_stacked_sel[[5]]$haswmhv2, 
           mat = TRUE) 

summary(imp_stacked_sel[[5]]$den_sel_f)
describeBy(imp_stacked_sel[[5]]$den_sel_f, 
           imp_stacked_sel[[5]]$haswmhv2, 
           mat = TRUE) 

summary(imp_stacked_sel[[20]]$sw_sel) 
describeBy(imp_stacked_sel[[20]]$sw_sel, 
           imp_stacked_sel[[20]]$haswmhv2, 
           mat = TRUE) 

bal.tab(haswmhv2 ~ 
          ethnicity +
          age_1 + # using baseline age 
          gender +
          # SES variables 
          edu + 
          income +
          hasnoinsurance + 
          socsupptot + 
          # proximal risk factors
          bmi +
          rapa_1_total + # physical act scores
          rapa_2_total + 
          smkever + 
          gds_total + 
          # lab values for cardiometabolic RF
          ldl + 
          gluc + 
          a1c + 
          egfr_nonaa + 
          egfr_aa + 
          choltot + 
          # clinical dx vars
          cdx_dep +
          cdx_dyslipid + 
          cdx_cvd +
          cdx_dm,
        data = imp_stacked_sel[[3]],
        weights = "sw_sel")

# cleaning up work space
rm(i, m, data, num_sel_wmhv, den_sel_wmhv)
```

# Creating total weights
For exposure weights, we will multiply exposure * selection weights. For the mediator weights, we include exposure, mediator, and selection weights. 
```{r IPW with all, echo = TRUE}
# Prepping some lists/variables for for-loop 
m <- max(imp_stacked$".imp") # number of imputed datasets
imp_stacked_wmhv_sbp <- list() # list for mediator weights 

# For-loop to obtain numerator model for exposure 
for (i in seq_along(1:m)){
  
  # extract one imputed dataset at a time
  data <- imp_stacked %>% filter(`.imp` == i)
  
  # Creating stabilized IPW for exposure
  imp_stacked_wmhv_sbp[[i]] <- data %>%
    
    # getting weights from the models above
    bind_cols(sw_exp = imp_stacked_exp[[i]]$sw_exp, 
              sw_med = imp_stacked_med[[i]]$sw_med,
              sw_sel = imp_stacked_sel[[i]]$sw_sel) %>% 
    
    # creating final exposure weights
    mutate(sw_exp_sel = (sw_exp*sw_sel)) %>% 
    
    ## truncating at 1st and 99th percentile 
    mutate(sw_exp_f = case_when(
      sw_exp_sel < quantile(sw_exp_sel, 0.01) ~ quantile(sw_exp_sel, 0.01),
      sw_exp_sel > quantile(sw_exp_sel, 0.99) ~ quantile(sw_exp_sel, 0.99),
      TRUE ~ sw_exp_sel)) %>% 
    
    # creating final mediator weights
    mutate(sw_med_sel = (sw_exp*sw_med*sw_sel)) %>% 
    
    ## truncating at 1st and 99th percentile 
    mutate(sw_med_f = case_when(
      sw_med_sel < quantile(sw_med_sel, 0.01) ~ quantile(sw_med_sel, 0.01),
      sw_med_sel > quantile(sw_med_sel, 0.99) ~ quantile(sw_med_sel, 0.99),
      TRUE ~ sw_med_sel)) 
  
}
```
## Examining final IPWs
Final exposure weights mean ~ 1, but max ranging 5s, will truncate at 1%tile and 99%tile (see above section for edits to code).

Final exposure weights mean ~ranging 1-3, max in the 1000s, will truncate (see above section for edits to code).
```{r check IPW all, echo = TRUE}
# switch index to check all datasets
summary(imp_stacked_wmhv_sbp[[20]]$sw_exp_f) 
describeBy(imp_stacked_wmhv_sbp[[20]]$sw_exp_f, 
           imp_stacked_wmhv_sbp[[20]]$ethnicity, 
           mat = TRUE) 

summary(imp_stacked_wmhv_sbp[[20]]$sw_med_f) 
describeBy(imp_stacked_wmhv_sbp[[20]]$sw_med_f, 
           imp_stacked_wmhv_sbp[[20]]$ethnicity, 
           mat = TRUE) 

# for exposure weights 
bal.tab(as.factor(ethnicity) ~
          age_1 + # using baseline age 
          gender +
          # SES variables 
          edu + 
          income +
          hasnoinsurance + 
          socsupptot + 
          # proximal risk factors
          bmi +
          rapa_1_total + # physical act scores
          rapa_2_total + 
          smkever + 
          gds_total + 
          # lab values for cardiometabolic RF
          ldl + 
          gluc + 
          a1c + 
          egfr_nonaa + 
          egfr_aa + 
          choltot + 
          # clinical dx vars
          cdx_dep +
          cdx_dyslipid + 
          cdx_cvd +
          cdx_dm,
        data = imp_stacked_wmhv_sbp[[5]],
        weights = "sw_med_f")

# visualizing balance before and after weighting for exposure weights
## For exposure weights 
exp_cov <- do.call(rbind, imp_stacked_wmhv_sbp) %>% 
  select(c("age_1", "ethnicity", "gender", "edu", "income", "hasnoinsurance", "socsupptot", "bmi", "rapa_1_total", "rapa_2_total", "smkever", "gds_total", "ldl", "gluc", "a1c", "egfr_nonaa", "egfr_aa", "choltot", "cdx_dep", "cdx_dyslipid", "cdx_cvd", "cdx_dm", "sw_exp_f"))

cov_names <- data.frame(
  old = c("age_1", "ethnicity", "gender", "edu", "income", "hasnoinsurance", "socsupptot", "bmi", "rapa_1_total", "rapa_2_total", "smkever", "gds_total", "ldl", "gluc", "a1c", "egfr_nonaa", "egfr_aa", "choltot", "cdx_dep", "cdx_dyslipid", "cdx_cvd", "cdx_dm", "sw_exp_f"),
  new = c("Age (Years)", "Race/Ethnicity", "Sex/Gender", "Years of Education", "Income", "Has No Insurance", "Social Support Score", "BMI", "Physical Activity Score (Aerobic)", "Physical Activity Score (Strength)", "Ever Smoked", "GDS Score", "LDL-C", "Glucose", "A1C", "EGFR-nonAA", "EGFR-AA", "Total Cholesterol", "Depression", "Dyslipidemia", "CVD", "Diabetes", "IPW for Exposure") ) 

love.plot(as.factor(ethnicity) ~
          age_1 + # using baseline age 
          gender +
          # SES variables 
          edu + 
          income +
          hasnoinsurance + 
          socsupptot + 
          # proximal risk factors
          bmi +
          rapa_1_total + # physical act scores
          rapa_2_total + 
          smkever + 
          gds_total + 
          # lab values for cardiometabolic RF
          ldl + 
          gluc + 
          a1c + 
          egfr_nonaa + 
          egfr_aa + 
          choltot + 
          # clinical dx vars
          cdx_dep +
          cdx_dyslipid + 
          cdx_cvd +
          cdx_dm,
          weights = "sw_exp_f",
          data = exp_cov,
          threshold = c(m=0.2),
          var.names = cov_names,
          title = "Covariate Balance for SBP Exposure Weights (WMHV)") -> lp_exp_wmhv_sbp

# for mediator weights
bal.tab(as.factor(has_htn) ~
          age_1 + # using baseline age 
          gender +
          # SES variables 
          edu + 
          income +
          hasnoinsurance + 
          socsupptot + 
          # proximal risk factors
          bmi +
          rapa_1_total + # physical act scores
          rapa_2_total + 
          smkever + 
          gds_total + 
          sbpavg + 
          dbpavg + 
          # lab values for cardiometabolic RF
          ldl + 
          gluc + 
          a1c + 
          egfr_nonaa + 
          egfr_aa + 
          choltot + 
          # clinical dx vars
          cdx_dep +
          cdx_dyslipid + 
          cdx_cvd +
          cdx_dm,
        data = imp_stacked_wmhv_sbp[[5]],
        weights = "sw_med_f")

## for mediator weights
med_cov <- do.call(rbind, imp_stacked_wmhv_sbp) %>% 
  select(c("age_1", "has_htn", "ethnicity", "gender", "edu", "income", "hasnoinsurance", "socsupptot", "bmi", "rapa_1_total", "rapa_2_total", "smkever", "gds_total", "ldl", "gluc", "a1c", "egfr_nonaa", "egfr_aa", "choltot", "cdx_dep", "cdx_dyslipid", "cdx_cvd", "cdx_dm", "sw_med_f"))

cov_names <- data.frame(
  old = c("age_1", "has_htn", "ethnicity", "gender", "edu", "income", "hasnoinsurance", "socsupptot", "bmi", "rapa_1_total", "rapa_2_total", "smkever", "gds_total", "ldl", "gluc", "a1c", "egfr_nonaa", "egfr_aa", "choltot", "cdx_dep", "cdx_dyslipid", "cdx_cvd", "cdx_dm", "sw_med_f"),
  new = c("Age (Years)", "Has HTN", "Race/Ethnicity", "Sex/Gender", "Years of Education", "Income", "Has No Insurance", "Social Support Score", "BMI", "Physical Activity Score (Aerobic)", "Physical Activity Score (Strength)", "Ever Smoked", "GDS Score", "LDL-C", "Glucose", "A1C", "EGFR-nonAA", "EGFR-AA", "Total Cholesterol", "Depression", "Dyslipidemia", "CVD", "Diabetes", "IPW for Mediator") ) 

love.plot(as.factor(has_htn) ~
          age_1 + # using baseline age 
          gender +
          ethnicity + 
          # SES variables 
          edu + 
          income +
          hasnoinsurance + 
          socsupptot + 
          # proximal risk factors
          bmi +
          rapa_1_total + # physical act scores
          rapa_2_total + 
          smkever + 
          gds_total + 
          # lab values for cardiometabolic RF
          ldl + 
          gluc + 
          a1c + 
          egfr_nonaa + 
          egfr_aa + 
          choltot + 
          # clinical dx vars
          cdx_dep +
          cdx_dyslipid + 
          cdx_cvd +
          cdx_dm,
          weights = "sw_med_f",
          data = med_cov,
          threshold = c(m=0.2),
          var.names = cov_names,
          title = "Covariate Balance for SBP Mediator Weights (WMHV)") -> lp_med_wmhv_sbp

# saving figures for love.plot for covariate balance
## exported through viewer

# table of weights per imputed dataset
do.call(rbind, imp_stacked_wmhv_sbp) %>%
  as.data.frame() %>% 
  group_by(`.imp`) %>% 
  summarise(mean_expIPW = mean(sw_exp_f),
            sd_expIPW = sd(sw_exp_f),
            min_expIPW = min(sw_exp_f),
            max_expIPW = max(sw_exp_f),
            mean_medIPW = mean(sw_med_f),
            sd_medIPW = sd(sw_med_f),
            min_medIPW = min(sw_med_f),
            max_medIPW = max(sw_med_f)) %>%
  rename("Mean (Exposure IPW)" = mean_expIPW,
         "SD (Exposure IPW)" = sd_expIPW,
         "Min (Exposure IPW)" = min_expIPW,
         "Max (Exposure IPW)" = max_expIPW,
         "Mean (Mediator IPW)" = mean_medIPW,
         "SD (Mediator IPW)" = sd_medIPW,
         "Min (Mediator IPW)" = min_medIPW,
         "Max (Mediator IPW)" = max_medIPW,
         "Imputation" = `.imp`) %>% 
  mutate_if(is.numeric, round, 2) %>% 
  gt() -> tbl_wgts_wmhv_sbp

# saving tables
gtsave(tbl_wgts_wmhv_sbp, file = "./results/tbl_wgts_wmhv_sbp.docx")

# saving dataset for analysis 
save(imp_stacked_wmhv_sbp, file = "./analysis data/imp_stacked_wmhv_sbp.Rdata")

# cleaning up workspace
rm(i, m, data, imp_stacked_exp, imp_stacked_med, imp_stacked_sel, cov_names, exp_cov, lp_exp_wmhv_sbp, lp_med_wmhv_sbp, med_cov)
```
# Estimating TE and CDEs for WMHV 
## Total effect
For total effect, we can model logWMHV_2 as a function of race/ethnicity, weighted by the final exposure IPW (which includes the IPW for selection into the measurement of WMHV at visit 2).
```{r te wmhv, echo = TRUE}
# loading stacked imputed data 
load("./analysis data/imp_stacked_wmhv_sbp.Rdata") 

# creating survey design in order to use weights
m <- length(imp_stacked_wmhv_sbp) # number of imputed datasets
te_wmhv <- list() # list for mediator weights 

# For-loop to estimate MSM for total effect of race/ethnicity on logWMHV
for (i in seq_along(1:m)){
  
  # creating survey design to implement weights 
  design_te_wmhv <- svydesign(ids = ~1, 
                              weights = ~sw_exp_f, 
                              data = imp_stacked_wmhv_sbp[[i]])
  
  # estimating total effect of race/ethnicity on log-transformed WMHV 
  te_wmhv[[i]] <- svyglm(logwmhv_2 ~
                           as.factor(ethnicity) +
                           as.factor(mriscanner_2) + 
                           scale(icv_1), # correcting for ICV
                         design = design_te_wmhv) 
}

# Pooling estimates across imputations 
pool_te_wmhv <- as.mira(te_wmhv) %>% pool()
summary(pool_te_wmhv)

# Making results presentable 
te_wmhv_res <- summary(pool_te_wmhv, conf.int = TRUE) %>%
  as.data.frame() %>% 
  mutate("Estimate" = exp(estimate)) %>% 
  mutate("LCL" = exp(conf.low)) %>% 
  mutate("UCL" = exp(conf.high)) %>%
  rename("p-value" = p.value) %>% 
  mutate(across(is.numeric, round, 2)) %>%
  mutate("Group" = case_when(
    term == "as.factor(ethnicity)Black" ~ "Black",
    term == "as.factor(ethnicity)Hispanic" ~ "Hispanic")) %>% 
  select(Group, Estimate, LCL, UCL, "p-value") %>% 
  filter(Group %in% c("Black", "Hispanic")) 

te_wmhv_res %>% gt()

# cleaning up workspace
rm(i, m, design_te_wmhv, te_wmhv)

```
## Controlled direct effect
For the CDE, we model logWMHV_2 as a function of the race/ethnicity, hypertension, and their interaction, weighted by the SW as calculated by the cumulative product of both the exposure, mediator, and selection weights. 
```{r cde wmhv, echo = TRUE}
# creating survey design in order to use weights
m <- length(imp_stacked_wmhv_sbp)# number of imputed datasets
cde_sbp_nhw <- list() # list for results using SBP centered on NHW mean
cde_sbp_aha <- list() # list for results using SBP centered on ACC/AHA SBP goal of <120 as "normal" 
imp_stacked_wmhv_sbp2 <- list() # for new datasets created in the loop

# For-loop to estimate MSM for total effect of race/ethnicity on logWMHV
for (i in seq_along(1:m)){
  
  # calculate mean SBP in NHW
  imp_stacked_wmhv_sbp[[i]] %>% group_by(ethnicity) %>%
    summarize_at(vars(sbpavg), list(mean_SBP = mean)) %>%
    filter(ethnicity == "White") %>% select(mean_SBP) -> sbp_nhw
  
  # creating SBP variable centered on mean for NHW, i.e. making the mean SBP for NHW the reference value
  # creating SBP variable centered on AHA goal of SBP<120
  imp_stacked_wmhv_sbp2[[i]] <- imp_stacked_wmhv_sbp[[i]] %>% 
    mutate(sbp_ref_nhw = sbpavg-(sbp_nhw$mean_SBP)) %>%
    mutate(sbp_ref_aha = sbpavg-120)
  
  # creating survey design to implement weights
  design_cde_wmhv <- svydesign(ids = ~1, 
                               weights = ~sw_med_f, 
                               data = imp_stacked_wmhv_sbp2[[i]])
  
  # estimating CDE of race/ethnicity on log-transformed WMHV, with avg SBP centered on the mean for NHW
  cde_sbp_nhw[[i]] <- svyglm(logwmhv_2 ~
                               as.factor(ethnicity)*sbp_ref_nhw +
                               as.factor(mriscanner_2) + 
                               scale(icv_1),
                              design = design_cde_wmhv)
  
  # estimating CDE of race/ethnicity on log-transformed WMHV, with avg SBP centered on the AHA goal of SBP<120
  cde_sbp_aha[[i]] <- svyglm(logwmhv_2 ~
                               as.factor(ethnicity)*sbp_ref_aha +
                               as.factor(mriscanner_2) + 
                               scale(icv_1),
                              design = design_cde_wmhv)
  
}

# Pooling estimates across imputations 
pool_cde_nhw <- as.mira(cde_sbp_nhw) %>% pool()
summary(pool_cde_nhw)

pool_cde_aha <- as.mira(cde_sbp_aha) %>% pool()
summary(cde_sbp_aha)

# Making results presentable 
## For SBP held at NHW mean 
cde_nhw_res <- summary(pool_cde_nhw, conf.int = TRUE) %>%
  as.data.frame() %>% 
  mutate("Estimate" = exp(estimate)) %>% 
  mutate("LCL" = exp(conf.low)) %>% 
  mutate("UCL" = exp(conf.high)) %>%
  rename("p-value" = p.value) %>% 
  mutate(across(is.numeric, round, 2)) %>%
  mutate("Group" = case_when(
    term == "(Intercept)" ~ "Intercept",
    term == "as.factor(ethnicity)Black" ~ "Black",
    term == "as.factor(ethnicity)Hispanic" ~ "Hispanic",
    term == "sbp_ref_nhw" ~ "Systolic Blood Pressure (mmHg), centered at mean for NHW",
    term == "as.factor(ethnicity)Black:sbp_ref_nhw" ~ "Black*SBP",
    term == "as.factor(ethnicity)Hispanic:sbp_ref_nhw" ~ "Hispanic*SBP",
    term == "scale(icv_1)" ~ "ICV"))  %>% 
  select(Group, Estimate, LCL, UCL, "p-value") %>% 
  filter(Group %in% c("Black", "Hispanic")) 

cde_nhw_res %>% gt()

## For SBP held at NHW mean 
cde_aha_res <- summary(pool_cde_aha, conf.int = TRUE) %>%
  as.data.frame() %>% 
  mutate("Estimate" = exp(estimate)) %>% 
  mutate("LCL" = exp(conf.low)) %>% 
  mutate("UCL" = exp(conf.high)) %>%
  rename("p-value" = p.value) %>% 
  mutate(across(is.numeric, round, 2)) %>%
  mutate("Group" = case_when(
    term == "(Intercept)" ~ "Intercept",
    term == "as.factor(ethnicity)Black" ~ "Black",
    term == "as.factor(ethnicity)Hispanic" ~ "Hispanic",
    term == "sbp_ref_aha" ~ "Systolic Blood Pressure (mmHg), centered at SBP=120 per AHA",
    term == "as.factor(ethnicity)Black:sbp_ref_aha" ~ "Black*SBP",
    term == "as.factor(ethnicity)Hispanic:sbp_ref_aha" ~ "Hispanic*SBP",
    term == "scale(icv_1)" ~ "ICV"))  %>% 
  select(Group, Estimate, LCL, UCL, "p-value") %>% 
  filter(Group %in% c("Black", "Hispanic")) 

cde_aha_res %>% gt()

# cleaning up workspace
rm(i, m, data, design_cde_wmhv, cde_wmhv)
```

# Outputting WMHV results 
```{r res wmhv, echo = TRUE}
wmhv_res_sbp <- bind_rows(te_wmhv_res,
                          cde_nhw_res,
                          cde_aha_res) %>% 
  gt() %>% 
  tab_row_group(
    label = "Controlled Direct Effects of Race/Ethnicity on WMHV, Setting Mean SBP=120 mmHg",
    rows = 5:6) %>% 
  tab_row_group(
    label = "Controlled Direct Effects of Race/Ethnicity on WMHV, Setting Mean SBP as Mean SBP in NHW",
    rows = 3:4) %>% 
  tab_row_group(
    label = "Total Effects of Race/Ethnicity on WMHV",
    rows = 1:2) 
  
wmhv_res_sbp

save(wmhv_res_sbp, file = "./results/wmhv_res_sbp") # saving gt version, in case we want to merge gt
wmhv_res_sbp %>% gtsave("./results/wmhv_res_sbp.docx")
```
